## Background

This repository contains all the code and outputs used to recreate figure from the Schaefer et al. (2000) study.  Their research assessed stream water quality at four sites, in Bisley, Puerto Rico, in the years flanking the 1989 Hurricane Hugo. To determine stream water quality, they analyzed a 9-wk moving average of calcium, potassium, magnesium, nitrate, and ammonium. 

## Data

Raw data from the Schaefer et al. (2000) study was obtained from the Environmental Data Initiative Data Portal.
<https://doi.org/10.6073/PASTA/F31349BEBDC304F758718F4798D25458>

## Methods

To recreate the figure the data was processed and a function was applied to calculate the 9-week moving average of ion concentration. A line graph was used to visulaize the results from 1986 to 1994.


## Results

The figure shows the results of the 9-week moving average analysis of ion concentration sampled from 1986 to 1994 at four stream sites in Bisley, Puerto Rico.

```{r}
#| include: false
library(tidyverse)
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-height: 9
#| fig-width: 7
#| fig-align: center
# Visualizing analysis on stacked graphs, each ion is identified by line type
bpr_ma_long <- read_csv("../output/bpr_ma_long.csv")

ggplot(
  data = bpr_ma_long,
  mapping = aes(
    x = window_start,
    y = `Concentration (mg/L)`,
    linetype = site
  )
) +
  geom_line() +
  labs(
    x = "Year",
    y = "Ion Concentration (mg/L)",
    linetype = "Bisley, Puerto Rico
    Sample Site"
  ) +
  facet_wrap(~Ion, scales = "free", ncol = 1, strip.position = "left") +
  scale_linetype_manual(values = c("solid", "dashed", "dotted", "dotdash")) +
  geom_vline(xintercept = as.Date("1989-9-10"), linetype = "dashed") +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5),
    panel.grid = element_blank(),
    strip.background = element_blank(),
    legend.title = element_blank(),
    legend.position = "inside",
    legend.position.inside = c(0.9, 0.3),
    legend.background = element_blank()
  )
```

## Collaborators

Project was peer reviewed by Natasha Clemencon and Sarah Busby.

## Reference

McDowell, William H., and USDA Forest Service. International Institute Of Tropical Forestry (IITF). 2024. “Chemistry of Stream Water from the Luquillo Mountains.” Environmental Data Initiative. <https://doi.org/10.6073/PASTA/F31349BEBDC304F758718F4798D25458>

Schaefer, Douglas. A., William H. McDowell, Fredrick N. Scatena, and Clyde E. Asbury. 2000. “Effects of Hurricane Disturbance on Stream Water Concentrations and Fluxes in Eight Tropical Forest Watersheds of the Luquillo Experimental Forest, Puerto Rico.” Journal of Tropical Ecology 16 (2): 189–207. <https://doi.org/10.1017/s0266467400001358>